\documentclass[12pt]{article}
 \usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts}
 
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
 
\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
%If you want to title your bold things something different just make another thing exactly like this but replace "problem" with the name of the thing you want, like theorem or lemma or whatever
 
\usepackage{mathtools}

\input{style}

\newcommand{\lecture}[1]{\details{#1}}
\usepackage{graphicx}
\usepackage{mathtools}	
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{quest}[thm]{Question}

\theoremstyle{definition}
\theoremstyle{lemma}
\theoremstyle{conclusion}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{question}[theorem]{Question}

\def\zint{\mathbb{Z}}
\def\zp{\zint_{p}}
\def\qp{\mathbb{Q}_{p}}

\def\zmodnz#1{\zint/#1\zint}
\def\zmodpz{\zmodnz p}
\def\zmodpmz#1{\zmodnz{p^{#1}}}
\def\zmodpnz{\zmodnz{p^{n}}}
\def\gl#1{GL(2, #1)}
\def\glzp{\gl{\zp}}
\def\glqp{\gl{\qp}}
\def\matr#1 #2 #3 #4{
\
  \begin{pmatrix}
    {#1} & {#2}  \\
    {#3} & {#4} 
  \end{pmatrix}
\
}


\makeatletter
\let\c@equation\c@thm
\makeatother
\numberwithin{equation}{section}


\begin{document}
%\renewcommand{\qedsymbol}{\filledbox}
%Good resources for looking up how to do stuff:
%Binary operators: http://www.access2science.com/latex/Binary.html
%General help: http://en.wikibooks.org/wiki/LaTeX/Mathematics
%Or just google stuff
 
\title{Representation of P-adic Groups}
\author{Noam Peri, 204686034}
\maketitle

I will solve questions from the book [GH]
\begin{problem}{6.2}
Define $red_n : \glzp \rightarrow \gl {\zmodpnz}$ by 
$\matr a b c d \mapsto  \matr \overline{a} \overline{b} \overline{c} {\overline{d}}$ 
\\ where $a \mapsto \overline{a}$ is the natural transformation "mod $p^n$
\end{problem}
\begin{enumerate}

\item prove that $red_n$ is well defined
\begin{proof}
We will use the following claim:
\begin{claim} Let R be a commutative ring with a unit and $M \in M_{n \times n}(R)$. Then M is invertible iff det(M) is invertible in R.
\end{claim} % TODO: address the proof%
Notice that in both $\zmodpnz$ and in $\zp$ an element x is invertible iff $p \nmid x$
\\Finally note that for any matrix M in either $\glzp$ or $\gl {\zmodpnz}$
\begin{multline*}
\matr a b c d = \matr {a_0 + pa'} {b_0 + pb'} {c_0 + pc'} {d_0 + pd'} \text{is invertible}  \iff p \nmid det\matr {a_0 + pa'} {b_0 + pb'} {c_0 + pc'} {d_0 + pd'} \\ \iff p \nmid [det \matr {a_0} {b_0} {c_0} {d_0} + p \cdot A] \iff 
 p \nmid det\matr {a_0} {b_0} {c_0} {d_0} = det(red_1(M))
\end{multline*}
\\ And thus we can conclude that M is invertible iff $red_1(M)$ is invertible iff $red_n(M)$ is invertible.
\end{proof}
\item prove that $red_n$ is onto $\gl {\zmodpnz}$ with kernel $K_n$
\begin{proof}
Denote by  $i : \zmodpnz \xhookrightarrow {} \zp$ the natural inclusion morphism. 
We can expand it to $i^*: \glzp \rightarrow \gl {\zmodpnz}$
\\ Then its clear that $red_n \circ i^* = id_{\gl {\zmodpnz}}$ and thus $red_n$ must be onto $\gl {\zmodpnz}$. Clearly, 
\begin{multline*}
ker (red_n) = \{\matr a b c d | \overline{a}=1, \overline{b}=0, \overline{c}=0,  \overline{d}=1\} \\
= \{\matr {1 + p^na'} {0 + p^nb'} {0 + p^nc'} {1 + p^nd'} | a',b',c',d' \in \zp \} = K_n
\end{multline*}
\end{proof}
\item Calculate the index $[\glzp:K_n]$
\begin{proof}
Using the last results and the  homomorphism theorem, we can see that:
\\$\glzp/K_n \cong \gl {\zmodpnz}$ 
\\ We will count the size of the latter expression. Every matrix $M \in \gl {\zmodpnz}$ can be written as $M = \matr a b c d + p \cdot \matr {a'} {b'} {c'} {d'}$ where $a,b,c,d \in [p-1]$ and $a',b',c',d' \in [p^{n-1}-1]$. Using the observations from the first part, M is invertible iff the first matrix is invertible in $\glzp$.
And thus $|\gl {\zmodpnz}| = p^{4(n-1)}\cdot |\gl {\zmodpz}|$.\\ 
It easy to calculate $|\gl {\zmodpz}| = (p^2 - 1)(p^2 - p)$ (Initially choosing a non zero vector, and then choosing another one that is not proportional to the first one). 
\\Concluding that $[\glzp:K_n] = p^{4(n-1)} (p^2 - 1)(p^2 - p)$
\end{proof}
\end{enumerate}

\begin{problem}{6.3}
For a $n > m$ show that $K_n \unlhd K_m$ and calculate $|K_m/K_n|$.
\end{problem}

\begin{proof}
From the definition $K_n \leqslant K_m \leqslant \glzp$. Using Problem 6.2 $K_n \lhd \glzp $ (As the kernel of a group homomorphism). And so $K_n \unlhd K_m \unlhd \glzp$. 
\\ Using the third homomorphism theorems we know that: 
$$[\glzp : K_m]\cdot[K_m:K_n] = [\glzp:K_n]$$
Using the result from Problem 6.2 we obtain:
$$[K_m:K_n] = \frac{[\glzp:K_n]}{[\glzp : K_m]} = \frac{p^{4(n-1)} (p^2 - 1)(p^2 - p)}{p^{4(m-1)} (p^2 - 1)(p^2 - p)} = p^{4(n-m)}$$
\end{proof}

\begin{problem}{6.5}
Let $(\pi, V)$ be a representation of $\glqp$. Show that the following 2 definitions for smoothness are equivalent:
\begin{itemize}
\item[a.] $\forall v \in V$ the function:
 $T_v: \gl \qp \rightarrow V$, defined by: $T_v(g) = \pi(g)v$, is locally constant.
\item[b.] $\forall v \in V \textbf{ } \exists n \in \mathbb{N}^+ \textbf{ }s.t \textbf{ }\forall k \in K_n \textbf{ }\pi(k)v=v$
\end{itemize} 

\end{problem}

\begin{proof}  \hfill
\\ \underline{\textbf{a $\Rightarrow$ b}}: For every $v \in V$, we know that: $T_v(I_2) = \pi(I_2)v = v$. Due to the fact that $T_v$ is locally constant, there is a neighborhood $U$ of $I_2$ such that: $\forall g \in U \textbf{ } T_v(g) = v$
Using Problem 6.2, we know that $\{K_n\}_{n\geq0}$ is a basis for the open neighborhood of $I_2$, and thus there is $n \in \mathbb{N}$ such that $I_2 \in K_n \subseteq U $. Concluding that $\forall g \in K_n \textbf{ } T_v(g) = v$, as needed.
\\ 
\\ \underline{\textbf{b $\Rightarrow$ a}}: Let $g \in \glzp$, $v \in V$, there exists $n>0$ such that: $\forall k \in K_n \textbf{ } T_v(k) = v$. 
\\ $U = g \cdot K_n$ is an open neighborhood of g(as multiplying by g is a open transformation).
\\ Every element $\varphi \in g \cdot K_n$ can be written as $\varphi = g \cdot k$ for some $k \in K_n$.
$$T_v(\varphi)=T_v(g \cdot k)=\pi(g\cdot k)v = \pi(g)[\pi(k)v]=\pi(g)v=T_v(g)$$ and  $T_v$ is constant on $U$.
\end{proof}

\begin{problem}{6.6}
Let $(\pi, V)$ be a smooth representation of $\glqp$
\end{problem}
\begin{enumerate}
\item Prove that $V^{K_m} \subseteq V^{K_n}$ for $m \leq n$, and that $\bigcup_{n \geq 1} V^{K_n} = V$
\begin{proof}
Let $m \leq n$ then $K_n \subseteq K_m$ as so:
$$ V^{K_m} = \{v \in V | \forall k \in K_m \textbf{ } \pi(k)v=v\} \subseteq \{v \in V | \forall k \in K_n \textbf{ } \pi(k)v=v\} = V^{K_n}$$
In addition using definition b (from question 6.5) of smoothness, for every $v \in V$ there is $n \geq 1$ such that $v \in V^{K_n}$. 
\\Concluding that $V = V^{K_n}$
\end{proof}


\item Show that $V^{K_n}$ is $\glzp$-invariant for each integer $n \geq1$
\begin{proof}
In question 6.2 we've seen that $K_n \lhd \glzp$, than for every $k \in K_n$ and $g \in \glzp$ there is a $k' \in K_n$ such that $kg = gk'$.
\\Let $v \in V^{K_n}, g \in \glzp$, we will show that $\pi(g)v \in V^{K_n}$. Indeed for every $k \in K_n$: 
$$ \pi(k)[\pi(g)v] = \pi(k \cdot g) v = \pi(g \cdot k')v = \pi(g) [\pi(k')v] = \pi(g)v$$ 
\end{proof}
\item Conclude that if $(\pi, V)$ is admissible, then every $v \in V$ lies in a finite-dimensional $\glzp$-invariant subspace of V.
\begin{proof}
The admissibility of $(\pi, V)$ means that $V^{K_n}$ is finite-dimensional for every n. From part 1 and 2 we know that they are $\glzp$-invariant and cover V. 
\end{proof}
\end{enumerate}


\begin{problem}{6.22}
Suppose $\delta = \matr {p^{e_1}} 0 0 {p^{e_2}} $, where $e_1 \leq e_2$ are integers.
\end{problem}
\begin{enumerate}
\item  For each $n \geq 1$ prove the equality of indices:
$$[K_n:\delta^{-1}K_n\delta \cap K_n]=p^{e_2 - e_1} = [\delta^{-1}K_n\delta:\delta^{-1}K_n\delta \cap K_n]  $$
\begin{proof}
We start by showing the equality between the two indexes. To do so will use the following observations:  $\delta^t = \delta,\ (\delta^{-1})^{t} = \delta^{-1}, \ M \in K_n \text{ iff } M^t \in K_n$.
\\ We define an automorphism of $\glqp$:  $\varphi: M \mapsto \delta^{-1} M^t  \delta$. 
\\From the previous observations we can verify that $\varphi$ maps $K_n$ onto $\delta^{-1} K_n  \delta$ and vice versa. And thus it induces an isomorphism between them and we conclude that
$$[K_n:\delta^{-1}K_n\delta \cap K_n]=
[\varphi(K_n):\varphi(\delta^{-1}K_n\delta \cap K_n)] = [\delta^{-1}K_n\delta:\delta^{-1}K_n\delta \cap K_n]$$


Now, we will characterize the subgroup $\delta^{-1}K_n\delta \cap K_n$. 
\\ Fix $M= \matr a b c d \in K_n$, then $\delta M \delta^{-1} = \matr {a} {p^{e_1 - e_2} b} {p^{e_2 - e_1} c} {d}$. Recall that $e_2 - e_1 \geq 0$, and thus the requirements from $a, d, p^{e2 - e1}c$ are  fulfilled.   
$$M \in \delta^{-1}K_n\delta \Longleftrightarrow \delta M \delta^{-1} \in  K_n \Longleftrightarrow p^n | p^{e1 - e2}b \Longleftrightarrow p^{n + e_2 - e_1} | b$$
Now we can calculate the index. For any 
$M = \matr a b c d, N = \matr {\tilde{a}} {\tilde{b}} {\tilde{c}} {\tilde{d}} \in K_n$ are in the same coset of $\delta^{-1}K_n\delta \cap K_n$ iff $MN^{-1} \in  \delta^{-1}K_n\delta \cap K_n$.
N is clearly invertible and we can write $N^{-1} = \frac{1}{|N|} \matr {\tilde{d}} {-\tilde{b}} {-\tilde{c}} {\tilde{a}}$
\\Due to the fact that $K_n$ is a subgroup we only need to check that 
\\ $x := (MN^{-1})_{(1,2)} = \frac{1}{|N|}(-a\tilde{b} + b\tilde{a})$ satisfies  $p^{n + e_2 - e_1} |  x$ (in $\zp$).
$N \in K_n$ and so $|N|$ is invertible, we can than reduce the requirement to
$b\tilde{a}-a\tilde{b}\equiv 0 \mod p^{n + e_2 - e_1}$ ,and again using the fact that $\tilde{a}, a \equiv 1 \mod p$ and thus invertible modulo $ p^{n + e_2 - e_1}$,
we can finally conclude that M, N are in the same coset iff $$\frac{b}{a} \equiv \frac{\tilde{b}}{\tilde{a}}\  mod \ ( p^{n + e_2 - e_1})$$
And so the coset is determined by this ratio.
Knowing that $a \equiv 1, b \equiv 0  \mod p^n$, we can conclude that $\frac{b}{a} \equiv 0 \mod p^n$. And so the ratio is always part of the set $\{x \cdot p^n \mid x \in \{0, ..., p^{e_2- e_1} -1\}\}$. Choosing $a = 1, b = x\cdot p^n$ each element of this set is obtained. We can conclude that the index is exactly $p^{e_2- e_1}$. 
\end{proof}
\item Suppose $g \in \glqp$. prove the equality of indices: 
\\ $[K_n:g^{-1}K_ng \cap K_n] =  [g^{-1}K_n g :g^{-1}K_n g \cap K_n]$
\begin{proof}
Using the Cartan decomposition any such $g \in \glqp$ can be written as $g = x \delta y$ with $x,y \in \glzp$ and $\delta = \delta(e_1,e_2)$ as in part 1.
We define an autorphism $\psi: \glqp \rightarrow \glqp$ by: $\psi: M \mapsto y M y^{-1}$.
\\ $\psi(K_n) = K_n$ (as $y \in \glzp$, $K_n\lhd \glzp)$
\\ $\psi(g^{-1}K_ng) =y \cdot( y^{-1}\delta^{-1}x^{-1}K_n x \delta y) \cdot y^{-1} = \delta^{-1}(x^{-1}K_nx)\delta = \delta^{-1}K_n\delta$
\\ Using those observations:
\\ $[K_n:g^{-1}K_ng \cap K_n]\stackrel{\psi}{=}[K_n:\delta^{-1}K_n\delta \cap K_n] \stackrel{1}{=}  [\delta^{-1}K_n\delta:\delta^{-1}K_n\delta \cap K_n] \stackrel{\psi^{-1}}{=} [g^{-1}K_n g :g^{-1}K_n g \cap K_n]$
\end{proof}
\item Let I be the common index from part (2) prove that:
$$ \mu(K_n) = I \cdot \mu(g^{-1}K_ng \cap K_n)= \mu(g^{-1}K_ng)$$ for all $g \in \glqp$ Deduce that $\mu$ is right invariant.
\begin{proof}
First notice that $g^{-1}K_ng$ in an compact open subset of $\glqp$ as multiplying by an element from wither side is an continuous and open morphism. An intersection of open compact subsets is also compact and open, and thus $\mu(g^{-1} K_n g)$, $\mu(g^{-1} K_n g \cap K_n)$ are well defined.
\\ Using part 2, we know that both $K_n$ and $g^{-1} K_n g$  are a disjoint union of  I cosets of $ g^{-1} K_n g \cap K_n$ and thus $\mu(K_n) = I \mu( g^{-1} K_n g \cap K_n) = \mu( g^{-1} K_n g)$.
\\$\mu(K_n) = \mu(  g^{-1} K_n g) \stackrel{*}{=} \mu(g \cdot  g^{-1} K_n g) = \mu(K_n \cdot g)$ (we used the fact that $\mu$ is left invariant in *). And thus is right invariant.
\end{proof}
\end{enumerate}


\end{document}
